\bf Range Analysis for the IonMonkey JavaScript engine

### 2012-05-02 project-writings/paper.tex
'selected' range analysis equations since we also do things like charCodeAt and >>, << and shit

\caption{Range arithmetic definitions}
\label{fig:range_arith}
\end{figure}

@@

\begin{figure}[ht]
\begin{eqnarray*}
\Squash(n) &=& \begin{cases}
 -2^{31} &\text{if } n = -\infty \\
2^{31}-1 &\text{if } n = \infty \\
n&\text{otherwise } \\
\end{cases}\\
\Squash([l, h]) &=& [\Squash(l), \Squash(h)]
\end{eqnarray*}
%
\begin{eqnarray*}
x = \Phi(y, z) &\Rightarrow&
    \Range(x) = \Squash(\Range(y)) \cup \Squash(\Range(z)) \\
x = \beta(y, [l, h]) &\Rightarrow& \Range(x) = \Squash(\Range(y)) \cap [l, h] \\
x = y + z &\Rightarrow& \Range(x) = \Squash(\Range(y)) + \Squash(\Range(z)) \\
x = y - z &\Rightarrow& \Range(x) = \Squash(\Range(y)) - \Squash(\Range(z)) \\
x = y \x z &\Rightarrow& \Range(x) = \Squash(\Range(y)) \x \Squash(\Range(z)) \\
x = c &\Rightarrow& \Range(x) = [c, c]
\end{eqnarray*}
\caption{Range analysis equations}
\label{fig:range_anal}
\end{figure}

@@

We ran each benchmark suite on IonMonkey, both with and without our range
analysis pass, and measured the performance differences. The results are
summarized below:

@@

>>>>

\caption{Selected range arithmetic definitions}
\label{fig:range_arith}
\end{figure}

@@

\begin{figure}[ht]
\begin{eqnarray*}
\Squash(n) &=& \begin{cases}
 -2^{31} &\text{if } n = -\infty \\
2^{31}-1 &\text{if } n = \infty \\
n&\text{otherwise } \\
\end{cases}\\
\Squash([l, h]) &=& [\Squash(l), \Squash(h)]
\end{eqnarray*}
%
\begin{eqnarray*}
x = \Phi(y, z) &\Rightarrow&
    \Range(x) = \Squash(\Range(y)) \cup \Squash(\Range(z)) \\
x = \beta(y, [l, h]) &\Rightarrow& \Range(x) = \Squash(\Range(y)) \cap [l, h] \\
x = y + z &\Rightarrow& \Range(x) = \Squash(\Range(y)) + \Squash(\Range(z)) \\
x = y - z &\Rightarrow& \Range(x) = \Squash(\Range(y)) - \Squash(\Range(z)) \\
x = y \x z &\Rightarrow& \Range(x) = \Squash(\Range(y)) \x \Squash(\Range(z)) \\
x = c &\Rightarrow& \Range(x) = [c, c]
\end{eqnarray*}
\caption{Selected range analysis equations}
\label{fig:range_anal}
\end{figure}

@@

We ran each benchmark suite on IonMonkey, both with and without our range
analysis pass, and measured the performance differences. The results are
summarized in Figures \ref{fig:v8bench}, \ref{fig:sunspider}, and \ref{fig:kraken}.

@@

In order to get a different view of how well our optimization was
performing, we annotated IonMonkey to emit for each addition
instruction compiled whether or not range analysis proved that it
could not overflow, and ran this against all of the tests in Kraken,
Sunspider, and v8bench. The data from these tests are presented in Figure
\ref{fig:count} (tests that were found to have very few total add
instructions were not included).

We ran this experiment in two configurations: with and without
on-stack replacement (OSR) enabled. On-stack replacement is a
mechanism by which IonMonkey can generate code that allows entering a
function at a point other than the start; specifically, it allows
entering in the middle of a loop. OSR allows IonMonkey to switch to
JITted code in the middle of a loop that is discovered to be hot,
instead of needing to wait for the function it is in to finish
first. OSR's impact on how many overlow checks we can eliminate is
somewhat unpredictable. OSR being on introduces new edges to the
control flow graph that can result in less precise range
information. However, OSR also causes more code to be JITted,
increasing potential opportunities to find ranges. Both of these
effects come into play in different tests.

These results show that our range analysis is generally quite
successful in eliminating overflow checks from programs, but our
timing results indicate that this does yield that substantial of a
performance improvement. We believe that this is because the overhead
of an essentially never taken branch on modern hardware are quite
low. While on a microbenchmark we constructed overflow check
elimination resulted in a substantial (around 30\%) speedup, it appears that
this does not translate well to more realistic code.

\begin{figure*}[!ht]
\centering
\includegraphics[width=0.98\textwidth]{counts_graph.png}
\caption{Percentage of add overflow checks eliminated}
\label{fig:count}
\end{figure*}
